---
title: "Data Check for Participant "
author: "FETZ Cologne"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: 
  html_document:
    theme: united
    highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE, warning=FALSE}
# makes sure that all the libraries used are installed
packages <- c("tidyverse", "psychonetrics", "qgraph", "scales")
if (length(missing_pkgs <-
           setdiff(packages, rownames(installed.packages()))) > 0) {
  message("Installing missing package(s): ",
          paste(missing_pkgs, collapse = ", "))
  install.packages(missing_pkgs)
}
library(tidyverse)
library(psychonetrics)
library(qgraph)
library(scales)

set.seed(1)
```
```{r, include=FALSE,warning=FALSE}
# set which participnt to process
participant_code <- "UKK1VIZYQMPMOCIX" # participant code (in "")

# load data
data <- read_csv("Questionnaire_Abfrage.csv")
participant_ids <- read_csv("Participants_Study_Time_Range.csv")

# convert participant code to participant ID 
participant_id <- participant_ids %>%
  filter(Participant_Code == participant_code) %>%
  select(Participant_ID) %>%
  as.numeric()

```
## We're looking at participant `r participant_code` with data ID `r participant_id`.

```{r, echo=FALSE}
# name vars
label_vars = c(
  "Ich bin traurig",
  "Ich nehme Dinge wahr, die andere \n    Menschen nicht wahrnehmen können",
  "Ich habe Schwierigkeiten, \n    mich zu konzentrieren",
  "Ich bin kontaktfreudig",
  "Ich fühle mich gestresst",
  "Ich bin zufrieden mit mir",
  "Ich fühle mich ängstlich",
  "Es fällt mir schwer, \n    mich zu Dingen zu motivieren",
  "Ich bin misstrauisch \n    gegenüber anderen Menschen",
  "Ich grüble"
)
final_vars <- Vars <- paste0("var_", c(as.character(1:10)))

# data preprocessing & check
check_data <- data %>%
  mutate(date_ESM = as.Date(DateTime),
         time_ESM = DateTime) %>%
  filter(Participant_ID == participant_id) %>%
  mutate(
    Day_Study = date_ESM - .$date_ESM[1] + 1,
    DateTime = str_replace(DateTime, "UTC", "CET"),
    Query_of_Day = case_when(
      as.POSIXct(paste(date_ESM, "08:30:00")) <= DateTime &
        as.POSIXct(paste(date_ESM, "10:30:00")) >= DateTime ~ 1,
      as.POSIXct(paste(date_ESM, "11:30:00")) <= DateTime &
        as.POSIXct(paste(date_ESM, "13:30:00")) >= DateTime ~ 2,
      as.POSIXct(paste(date_ESM, "14:30:00")) <= DateTime &
        as.POSIXct(paste(date_ESM, "16:30:00")) >= DateTime ~ 3,
      as.POSIXct(paste(date_ESM, "17:30:00")) <= DateTime &
        as.POSIXct(paste(date_ESM, "19:30:00")) >= DateTime ~ 4,
      as.POSIXct(paste(date_ESM, "20:30:00")) <= DateTime &
        as.POSIXct(paste(date_ESM, "22:30:00")) >= DateTime ~ 5
    )
  )  %>%
  mutate(
    Lead_Day = lead(Query_of_Day, 1),
    Query_of_Day = case_when(
      Lead_Day == Query_of_Day ~ Query_of_Day - 1,
      is.na(Lead_Day == Query_of_Day) ~ Query_of_Day,
      TRUE ~ Query_of_Day
    )
  ) %>%
  mutate(
    Ref_Time = case_when(
      Query_of_Day == 1 ~ as.POSIXct(paste(date_ESM, "09:30:00")),
      Query_of_Day == 2  ~  as.POSIXct(paste(date_ESM, "12:30:00")),
      Query_of_Day == 3 ~ as.POSIXct(paste(date_ESM, "15:30:00")),
      Query_of_Day == 4 ~ as.POSIXct(paste(date_ESM, "18:30:00")),
      Query_of_Day == 5 ~ as.POSIXct(paste(date_ESM, "21:30:00"))
    )
  ) %>%
  mutate(Diff_Time = difftime(DateTime, Ref_Time,  units = "mins")) %>%
  mutate(Query_of_Day = ifelse(Diff_Time > 60, NA, Query_of_Day)) %>%
  rename_at(vars(matches("-1-0")), ~ final_vars) %>%
  mutate(across(matches("var_"), function(x)
    ifelse(is.na(.$Query_of_Day), NA, x))) %>%
  select(-c(Timestamp, Lead_Day, Ref_Time, Diff_Time))

# some stats for text
first_day <- min(check_data$date_ESM)
last_day <- max(check_data$date_ESM)
number_days <- max(check_data$Day_Study)
missings_of_filled_out <- sum(is.na(check_data$var_1))/nrow(check_data)*100
filled_out_total <- sum(!is.na(check_data$var_1))/min(14*5, as.numeric(number_days)*5)*100
filled_in_days <- length(unique(as.numeric(check_data$Day_Study)))

# severity plot
severity_plot <- data %>%
  mutate(date_ESM = as.Date(DateTime),
         time_ESM = DateTime) %>%
  filter(Participant_ID == participant_id) %>%
  rename_at(vars(matches("1-0")), ~ label_vars) %>%
  select(label_vars) %>%
  summarise_all(mean) %>%
  pivot_longer(cols = everything(),
               names_to = "Erlebnis",
               values_to = "Wert") %>%
  mutate(Erlebnis = reorder(Erlebnis, Wert)) %>%
  ggplot(., aes(x = Erlebnis, y = Wert, fill = Erlebnis)) +
  geom_bar(stat = "identity") +
  xlab("") +
  ylab("\n Durchschnittliche Stärke (%)") +
  coord_flip() + theme_classic() +
  theme(
    legend.position  = "none",
    axis.title = element_text(size = 14, colour = "black"),
    axis.text = element_text(size = 14, color = "black")
  ) +
  scale_fill_viridis_d()

most_variable_items <- data %>%
  mutate(date_ESM = as.Date(DateTime),
         time_ESM = DateTime) %>%
  filter(Participant_ID == participant_id) %>%
  rename_at(vars(matches("1-0")), ~ label_vars) %>%
  select(label_vars) %>%
  summarise_all(sd) %>%
  pivot_longer(cols = everything(),
               names_to = "Erlebnis",
               values_to = "SD") %>%
  top_n(10, SD) %>% arrange(-SD)

fluctuation_plot <- data %>%
  mutate(date_ESM = as.Date(DateTime),
         time_ESM = DateTime) %>%
  filter(Participant_ID == participant_id) %>%
  rename_at(vars(matches("1-0")), ~ label_vars) %>%
  select(most_variable_items$Erlebnis, time_ESM) %>%
  pivot_longer(cols = most_variable_items$Erlebnis,
               names_to = "Erlebnis",
               values_to = "Wert") %>%
  mutate(Erlebnis = factor(Erlebnis, levels = most_variable_items$Erlebnis)) %>%
  ggplot(., aes(x = time_ESM, y = Wert, color = Erlebnis)) + facet_wrap(~ Erlebnis,
                                                                          
                                                                                 ncol = 1) +  geom_line(size =
                                                                                                             1) +
  geom_smooth(method = "loess", formula = y ~ x) +
  theme_classic() +
  xlab("\n Datum") +
  theme(
    legend.position  = "none",
    strip.text = element_text(size = 14, colour = "black"),
    axis.title = element_text(size = 14, colour = "black"),
    axis.text = element_text(size = 14, colour = "black")
  ) +
  scale_color_viridis_d(direction = -1) +
  scale_y_continuous(limits = c(-10, 100),
                     breaks = c(0, 25, 50, 75, 100)) +
  scale_x_datetime(breaks =  date_breaks("2 days"), labels = date_format("%d.%m."))

```
## General Info
Participant `r participant_code` entered the PhenoNetz-Study on `r format(first_day, '%d %B, %Y')`. 
The last day in the study was `r format(last_day, '%d %B, %Y')`.
He/she completed `r number_days` days in the study, of which he/she filled out surveys on `r filled_in_days` days.

## Raw Data 
```{r, echo = F}
data %>% filter(Participant_ID == participant_id) %>% print(n=Inf, width=Inf)
```

## Preprocessed Data
```{r, echo = F}
check_data %>% print(n=Inf, width=Inf)
```


## Missingness / Valid Beeps
He/she completed a total of `r sum(!is.na(check_data$var_1))+sum(is.na(check_data$var_1))` surveys, of which `r sum(!is.na(check_data$var_1))` were valid. Participant `r participant_code` has `r round(missings_of_filled_out,2)` % missing values of those he/she filled out. Of all possible beeps (max. = 70), `r 100-round(filled_out_total,2)` % are missing.

## Experience Sampling Items
### Mean Endorsement
````{r, fig.width = 8, fig.height = 6}
severity_plot
```

## Fluctuation
Items sorted by how much the fluctuated:
````{r,  fig.width = 6, fig.height = 18}
most_variable_items
fluctuation_plot
```
