---
title: 'PhenoNetz: Group-Level Analyses'
date: "`r format(Sys.time(), '%d. %B %Y')`"
author: Linda Betz
fontsize: 8 pt
output: beamer_presentation
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=F, echo=F}
library(tidyverse)
library(psychonetrics)
library(mlVAR)


data <- read_csv("Questionnaire_Abfrage.csv")

# function to determine zero range
zeroRange <- function(x, tol = .Machine$double.eps ^ 0.5) {
  x <- x %>% na.omit()
  if (length(x) == 1) return(TRUE)
  x <- range(x) / mean(x)
  isTRUE(all.equal(x[1], x[2], tolerance = tol))
}

# function to remove linear trends
removeLinearTrends <- function(df) {
  df <- as.data.frame(df)
  for (v in seq_along(Vars)) {
    if(zeroRange(df[, Vars[v]]) == TRUE) {df[, Vars[v]] <- df[, Vars[v]]} else {
    lmResult <- lm(unlist(df[, Vars[v]]) ~ df$time_ESM)
    if (anova(lmResult)[["Pr(>F)"]][1] < .05) {
      df[, Vars[v]][!is.na(df[, Vars[v]])] <- residuals(lmResult)
    }
    }
  }
  return(as_tibble(df))
}

Participants_Study <- read_csv("Participants_Study_Time_Range.csv")

# define dropouts, irrelevant Test-IDs, etc.
Participants_Study %>%
  filter(
    !Participant_Code %in% c(
      "UKK1VIZYXFWIZJIM",
      "UKK1VIZYNEMXSMUK",
      "UKK1VIZYFOOKKBDR",
      "UKK1VIZYIAJHBQQD",
      "UKK1VIZYFTXUDQQV",
      "UKK1VIZYGJZWBENV",
      "UKK1VIZYTDGGOAFC",
      "UKK1VIZYCAATYTOJ",
      "UKK1VIZYPHWVDQPA",
      "UKK1VIZYTBLBGTBY"
    )
  ) %>% .$Participant_ID  -> final_participants

```

## Basic stats
Up to now, there have been `r data %>% filter(Participant_ID %in% final_participants) %>% distinct(Participant_ID) %>% nrow(.)` "non-dropout" participants in the study.

## Beeps per person
```{r}
final_vars <-
  Vars <-
  c(
    "var_1"  ,
    "var_2" ,
    "var_3",
    "var_4" ,
    "var_5" ,
    "var_6" ,
    "var_7"  ,
    "var_8",
    "var_9"  ,
    "var_10"
  )
# simple histogram
data %>% 
  filter(Participant_ID %in% final_participants) %>% 
  group_by(Participant_ID) %>%
  nest() %>%
  mutate(nrows = map_dbl(data, ~ nrow(.))) %>% ungroup() %>% select(nrows) -> beeps_per_person

ggplot(as_tibble(beeps_per_person), aes(x = nrows)) + geom_histogram(fill =
                                                                        	"#7070db",  binwidth = 10)  + theme_classic(base_size = 16) + ggtitle(label = paste0(
                                                                         "Median Number of Total Beeps (Total): ",
                                                                         median(beeps_per_person$nrows)
                                                                       )) +
  xlab("Total Beeps")
```


## Valid beeps
```{r}
data %>%
  filter(Participant_ID %in% final_participants) %>%
  mutate(date_ESM = as.Date(DateTime),
         time_ESM = DateTime) %>%
  mutate(
    DateTime = str_replace(DateTime, "UTC", "CET"),
    Query_of_Day = case_when(
      as.POSIXct(paste(date_ESM, "08:30:00")) <= DateTime &
        as.POSIXct(paste(date_ESM, "10:30:00")) >= DateTime ~ 1,
      as.POSIXct(paste(date_ESM, "11:30:00")) <= DateTime &
        as.POSIXct(paste(date_ESM, "13:30:00")) >= DateTime ~ 2,
      as.POSIXct(paste(date_ESM, "14:30:00")) <= DateTime &
        as.POSIXct(paste(date_ESM, "16:30:00")) >= DateTime ~ 3,
      as.POSIXct(paste(date_ESM, "17:30:00")) <= DateTime &
        as.POSIXct(paste(date_ESM, "19:30:00")) >= DateTime ~ 4,
      as.POSIXct(paste(date_ESM, "20:30:00")) <= DateTime &
        as.POSIXct(paste(date_ESM, "22:30:00")) >= DateTime ~ 5
    )
  )  %>%
  mutate(
    Lead_Day = lead(Query_of_Day, 1),
    Query_of_Day = case_when(
      Lead_Day == Query_of_Day ~ Query_of_Day - 1,
      is.na(Lead_Day == Query_of_Day) ~ Query_of_Day,
      TRUE ~ Query_of_Day
    )
  ) %>%
  mutate(
    Ref_Time = case_when(
      Query_of_Day == 1 ~ as.POSIXct(paste(date_ESM, "09:30:00")),
      Query_of_Day == 2 ~ as.POSIXct(paste(date_ESM, "12:30:00")),
      Query_of_Day == 3 ~ as.POSIXct(paste(date_ESM, "15:30:00")),
      Query_of_Day == 4 ~ as.POSIXct(paste(date_ESM, "18:30:00")),
      Query_of_Day == 5 ~ as.POSIXct(paste(date_ESM, "21:30:00"))
    )
  ) %>%
  mutate(Diff_Time = difftime(DateTime, Ref_Time,  units = "mins")) %>%
  mutate(Query_of_Day = ifelse(abs(Diff_Time) > 60, NA, Query_of_Day)) %>%
  rename_at(vars(matches("-1-0")), ~ final_vars) %>%
  mutate(across(matches("var_"), function(x)
    ifelse(is.na(.$Query_of_Day), NA, x))) %>%
  select(-c(Timestamp, Lead_Day, Ref_Time, Diff_Time)) %>%
  group_by(Participant_ID) %>%
  nest() %>%
  mutate(detrended_data = map(data, ~removeLinearTrends(.) )) %>%
  select(Participant_ID, detrended_data) %>% 
  unnest(cols = c(detrended_data)) %>% 
  ungroup() -> detrended_data


detrended_data %>% 
  group_by(Participant_ID) %>% 
  nest() %>% 
  mutate(nrows=map_dbl(data, ~sum(!is.na(.$Query_of_Day)))) %>% 
  ungroup() %>% 
  summarise(median = median(nrows)) -> median_valid_beeps

detrended_data %>% group_by(Participant_ID) %>%
  nest() %>%
  mutate(nrows = map_dbl(data, ~ sum(!is.na(.$Query_of_Day)))) %>%
  ggplot(., aes(x = nrows)) + geom_histogram(binwidth = 10, fill =  	 "#7070db")  + theme_classic(base_size =
                                                                                                   16) + ggtitle(label = paste0(
                                                                                                     "Median Number of Valid Beeps (Total): ",
                                                                                                     median_valid_beeps$median
                                                                                                   )) + xlab("Valid Beeps")
```

## Participants with less than 20 valid beeps
```{r, include=FALSE}
detrended_data %>%
  group_by(Participant_ID) %>%
  nest() %>%
  mutate(valid_beeps = map_int(data, ~sum(!is.na(.$Query_of_Day)))) %>%
  filter(valid_beeps >= 20) -> detrended_final_data

participants_total <- nrow(distinct(detrended_data, Participant_ID)) 

participants_enough_data <- nrow(distinct(detrended_final_data, Participant_ID)) 

median_beeps_in_excluded <- detrended_data %>%
  group_by(Participant_ID) %>%
  nest() %>%
  mutate(valid_beeps = map_int(data, ~sum(!is.na(.$Query_of_Day)))) %>%
  filter(valid_beeps < 20) %>% 
  ungroup() %>%
  summarise(median = median(valid_beeps))

```
After excluding dropouts/invalid IDs, there have been `r participants_total` participants total in the study. Of these, `r participants_enough_data` had at least 20 valid beeps. *These were used for all the following computations.*
Those participants that were excluded due to insufficient data had a median of `r median_beeps_in_excluded` valid beeps. Note that this includes participants who have less than 20 beeps because they have not yet finished the study.



## Valid beeps
```{r}
detrended_final_data %>%
  mutate(nrows = map_dbl(data, ~ sum(!is.na(.$Query_of_Day)))) %>%
  ungroup() %>%
  summarise(median = median(nrows)) -> median_valid_beeps

detrended_final_data %>%
  mutate(nrows = map_dbl(data, ~ sum(!is.na(.$Query_of_Day)))) %>%
  ggplot(., aes(x = nrows)) + geom_histogram(binwidth = 10, fill =  	 "#7070db")  + theme_classic(base_size =
                                                                                                   16) + ggtitle(label = paste0(
                                                                                                     "Median Number of Valid Beeps (Total): ",
                                                                                                     median_valid_beeps$median
                                                                                                   )) + xlab("Valid Beeps")
```

## Average number of total beeps per day
```{r}
detrended_final_data %>% 
  unnest(cols = c(data)) %>%
  group_by(Participant_ID) %>%
  mutate(date_ESM = as.Date(DateTime)) %>%
  nest() %>%
  mutate(per_day = map(data,
                       . %>% group_by(date_ESM) %>% summarise(n_per_day = n()))) %>% unnest(per_day) -> median_total

data %>% group_by(Participant_ID) %>%
  mutate(date_ESM = as.Date(DateTime)) %>%
  nest() %>%
  mutate(per_day = map(
    data,
    . %>% group_by(date_ESM) %>% summarise(n_per_day = n()) %>% summarise(n_per_day_median =
                                                                            median(n_per_day))
  )) %>% unnest(cols = c(data, per_day)) %>%
  distinct(Participant_ID , n_per_day_median) %>%
  ggplot(., aes(x = n_per_day_median)) + geom_histogram(binwidth = 0.5, fill =
                                                           	 "#7070db")  + ggtitle(label = paste0(
                                                            "Median Number of Beeps per Day: ",
                                                            median(median_total$n_per_day)
                                                          )) +
  theme_classic(base_size = 16) + xlab("Median of Beeps per Day")

```

## Loss of beeps (number)
```{r}
detrended_final_data %>%
  mutate(total_beeps = map_dbl(data, ~ nrow(.))) %>% select(Participant_ID, total_beeps) -> total_beeps

detrended_final_data %>%
  mutate(valid_beeps = map_dbl(data, ~ sum(!is.na(.$Query_of_Day)))) %>% select(Participant_ID, valid_beeps) -> valid_beeps

valid_beeps %>%
  left_join(total_beeps, by = "Participant_ID") %>%
  ungroup() %>%
  mutate(lost_beeps = total_beeps - valid_beeps) -> lost_beeps

ggplot(lost_beeps, aes(x = lost_beeps)) + geom_histogram(binwidth = 2, fill =
                                                            	 "#7070db")  + theme_classic(base_size = 16) +
  ggtitle(label = paste0("Median Number of Lost Beeps: ", median(lost_beeps$lost_beeps))) +
  xlab("Lost Beeps (#)")

```



## Loss of beeps (%)
```{r}
valid_beeps %>% 
  left_join(total_beeps, by = "Participant_ID") %>%
  mutate(lost_beeps = (1 - valid_beeps / total_beeps)) -> lost_beeps

ggplot(lost_beeps, aes(x = lost_beeps)) + geom_histogram(binwidth = 0.1, fill = 	 "#7070db")  + theme_classic(base_size=16) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  ggtitle(label = paste0("Median % of Lost Beeps: ", round(median(
    lost_beeps$lost_beeps
  ), 2) * 100)) +
  xlab("Lost Beeps (%)")
```


## Mean scores per item
```{r, fig.width=7.5, fig.height=6, warning=FALSE, echo=F, message=F}

labels_short <- c("traurig", "wahrnehmen", "konzentrieren", "kontaktfreudig", "gestresst", "zufrieden", "ängstlich", "motivieren", "misstrauisch", "grübeln")

data %>%
  filter(Participant_ID %in% detrended_final_data$Participant_ID) %>% 
  group_by(Participant_ID) %>%
  rename_at(vars(matches("-1-0")), ~ labels_short) %>%
  nest() %>%
  mutate(summarised_data = map(data,  . %>% summarise(across(labels_short, mean)))) %>%
  unnest(cols = c(summarised_data)) %>%
  pivot_longer(traurig:grübeln, values_to = "mean") %>%
   ggplot(aes(x = name,
             y = mean, fill = name)) +
  facet_wrap(~name, scales = "free") +
  geom_boxplot(alpha = 0.4) +
  geom_jitter(size = 0.6, alpha = 0.9) +
  ylab("Mean") +
  xlab("") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 12),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text = element_text(size = 12),
   legend.position = "none"
  ) + scale_fill_viridis_d()
```
```{r, include = FALSE, message = FALSE, warning= FALSE}
mlVAR(
  data = detrended_final_data %>% unnest(cols = c(data)),
  vars = final_vars,
  beepvar = "Questionnaire_of_Day",
  dayvar = "date_ESM",
  idvar = "Participant_ID"
) -> modelsmlvar
```
```{r}
centrality_temporal <- qgraph::centrality(getNet(modelsmlvar, "temporal"))
out_degree_temporal <- centrality_temporal$OutDegree %>% which.max(.)
in_degree_temporal <- centrality_temporal$InDegree %>% which.max(.)


centrality_contemp <- qgraph::centrality(getNet(modelsmlvar, "contemp"))
out_degree_contemp <- centrality_contemp$OutDegree %>% which.max(.)

```


## Temporal Network
The item most influenced *by other items* is "`r labels_short[in_degree_temporal]`". "`r labels_short[out_degree_temporal]`" influences *other items* most.
```{r,  message=F}
qgraph::qgraph(
  mlVAR::getNet(modelsmlvar, "temp", rule = "and"),
  layout = "spring",
  nodeNames = labels_short,
  labels = c(1:10),
  legend = T,
  theme = "Borkulo"
)
```


## Contemporaneous Network
The central item is "`r labels_short[out_degree_contemp]`".
```{r,  message=F}
qgraph::qgraph(
  mlVAR::getNet(modelsmlvar, "contemp", rule = "and"),
  layout = "spring",
  nodeNames = labels_short,
  labels = c(1:10),
  legend = T,
  theme = "Borkulo"
)
```